---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(bitstreamio)

if (FALSE) {
  # Check for functions which do not have an @examples block for roxygen
  system("grep -c examples man/*Rd", intern = TRUE) |> 
    grep(":0$", x = _, value = TRUE)
}
```

# bitstreamio

<!-- badges: start -->
![](https://img.shields.io/badge/cool-useless-green.svg)
[![CRAN](https://www.r-pkg.org/badges/version/bitstreamio)](https://CRAN.R-project.org/package=bitstreamio)
[![R-CMD-check](https://github.com/coolbutuseless/bitstreamio/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/coolbutuseless/bitstreamio-dev/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`bitstreamio` is a package for reading bits from a connection or raw vector.

This package has no requirement that reads/write operations be *byte aligned*.
This is useful for input/output with packed binary file formats e.g. h264-compressed
video, mp3 audio etc.

In addition to reading individual bits, this package also provides for:

* Reading/writing unaligned bytes as raw vectors
* Reading/writing unsigned integers (i.e. non-negative integers) at bit depths
  from 1 to 31
* Read/write signed/unsigned integers with [Exponential-Golomb coding ](https://en.wikipedia.org/wiki/Exponential-Golomb_coding)


## What's in the box

* Read write: bits, bytes, unsigned integers, integers coded with 
  exponential golomb (signed and unsigned variants).
    * `bs_write_bit()`, `bs_read_bit()`
    * `bs_write_byte()`, `bs_read_byte()`
    * `bs_write_uint()`, `bs_read_uint()`
    * `bs_write_uint_exp_golomb()`, `bs_read_uint_exp_golomb()`
    * `bs_write_sint_exp_golomb()`, `bs_read_sint_exp_golomb()`
* Bitstream utilities
    * `bs_open()`, `bs_close()`
    * `bs_is_aligned()`, `bs_align()`
    * `bs_advance()`, `bs_peek()`


## Installation

<!-- This package can be installed from CRAN -->

<!-- ``` r -->
<!-- install.packages('bitstreamio') -->
<!-- ``` -->

You can install the latest development version from
[GitHub](https://github.com/coolbutuseless/bitstreamio) with:

``` r
# install.package('remotes')
remotes::install_github('coolbutuseless/bitstreamio')
```

<!-- Pre-built source/binary versions can also be installed from -->
<!-- [R-universe](https://r-universe.dev) -->

<!-- ``` r -->
<!-- install.packages('bitstreamio', repos = c('https://coolbutuseless.r-universe.dev', 'https://cloud.r-project.org')) -->
<!-- ``` -->

## Read/Write with a raw vector

```{r example}
library(bitstreamio)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Write bits to a raw vector
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
bs <- bs_open(raw(), "w")
bs_write_bit(bs, c(TRUE , FALSE, FALSE, FALSE))
bs_write_byte(bs, c(0x01, 0x7f))  # write unaligned byte values
bs_write_bit(bs, c(FALSE, FALSE, FALSE, TRUE))

raw_vec <- bs_close(bs)
raw_vec


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Read bits back from raw vector
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
bs <- bs_open(raw_vec, mode = 'r')
bs_read_bit(bs, 4)
bs_read_byte(bs, 2) # unaligned byte read
bs_read_bit(bs, 4)
bs_close(bs)
```


## Reading/Write with a connection


```{r}
filename <- tempfile()

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Write bits to a file
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
con <- file(filename, "wb")
bs  <- bs_open(con, mode = 'w')

bs_write_bit(bs, c(TRUE , FALSE, FALSE, FALSE))
bs_write_byte(bs, c(1, 127))
bs_write_bit(bs, c(FALSE, FALSE, FALSE, TRUE))

bs_close(bs)
close(con)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Read the bits back in
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
con <- file(filename, "rb")
bs  <- bs_open(con, mode = 'r')

bs_read_bit(bs, 4)
bs_read_uint(bs, nbits = 8, n = 2)
bs_read_bit(bs, 4)

bs_close(bs)
close(con)
```


## Read/write unsigned integers at varying bit depths

```{r}
library(bitstreamio)

#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Write bits to a raw vector
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
bs <- bs_open(raw(), "w")
bs_write_uint(bs,  0:7 , nbits = 3) # write 8 * 3-bit integers
bs_write_uint(bs, 10:15, nbits = 4) # write 6 * 4-bit integers
raw_vec <- bs_close(bs)
raw_vec


#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
# Read bits back from raw vector
#~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
bs <- bs_open(raw_vec, mode = 'r')
bs_read_uint(bs, nbits = 3, n = 8)
bs_read_uint(bs, nbits = 4, n = 6)
bs_close(bs)
```





```{r echo=FALSE, eval=FALSE}
N <- 1000
set.seed(1)
raw_vec <- as.raw(sample(255, size = N, replace = TRUE))

f_ctypes <- function() {
  ctypesio::read_raw(raw_vec, n = N)
}

f_bitstreamio_bits <- function() {
  bs <- bs_open(raw_vec, 'r')
  on.exit(bs_close(bs))
  bs_read_bit(bs, N * 8)
}

f_bitstreamio_bytes <- function() {
  bs <- bs_open(raw_vec, 'r')
  on.exit(bs_close(bs))
  bs_read_byte(bs, N)
}


gc()
gc()
bench::mark(
  f_bitstreamio_bits(),
  f_bitstreamio_bytes(),
  f_ctypes(),
  check = FALSE
)

```









## Related Software

* [ctypesio](https://cran.r-project.org/package=ctypesio) - for reading/writing
  standard C types with connections e.g. `uint16`, `float`, `double` etc.
